---
title: "Risk Stratification"
author: 
date: 
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Risk stratification of diabetic patients for readmission

EDA and Feature engineering
```{r include=FALSE}
library(ggplot2)
library(stringr)
library(dplyr)
library(stringr)
library(caTools)
library(rpart)
library(caret)
library(car)
library(MASS)
```


```{r}
data <- read.csv("data/diabetic_data_original.csv", header=TRUE, 
                 na.strings = c("NA","na",""," ","?"), stringsAsFactors = FALSE)
```

Checking structure of data
```{r}
str(data)
```

Check for NA
```{r}
sapply(data, function(x) sum(is.na(x))) 
```

### Check variable
```{r}
# encounter_id
length(unique(data$encounter_id))
# 101766 variable, no duplicate
 #  remove encounterid
data <- subset(data, select=-c(encounter_id))
```


```{r}
# patient_nbr
# Save the patient id to identify the patien after analysis
Patient_Id <- data$patient_nbr
#  remove unwanted variable
data <- subset(data, select=-c(patient_nbr))
```

```{r}
# race
table(data$race)
# AfricanAmerican: 19210, Asian: 641, Caucasian: 76099, Hispanic: 2037, Other: 1506 
class(data$race)
data$race <- as.factor(data$race)
class(data$race)
```

```{r}
# gender
table(data$gender)
#Female : 54708, Male: 47055, Unknown/Invalid: 3
class(data$gender)
data$gender <- replace(data$gender, data$gender == "Unknown/Invalid", NA)
table(data$gender)
#convert gender to as.numeric
summary(factor(data$gender))
# convert to factor
data$gender <- as.factor(data$gender)
# give values 1 and 0 to the factors
levels(data$gender)<-c(1,0)
#convert to as.numeric
data$gender <- as.numeric(levels(data$gender))[data$gender]
summary(data$gender)
```

```{r}
# Age
table(data$age)
```

Reducing the levels
```{r include=FALSE}
data$age <- replace(data$age, data$age == "[0-10)", "1")
data$age <- replace(data$age, data$age == "[10-20)", "2")
data$age <- replace(data$age, data$age == "[20-30)","3")
data$age <- replace(data$age, data$age == "[30-40)","3")
data$age <- replace(data$age, data$age == "[40-50)", "4")
data$age <- replace(data$age, data$age == "[50-60)", "4")
data$age <- replace(data$age, data$age == "[60-70)", "5")
data$age <- replace(data$age, data$age == "[70-80)", "6")
data$age <- replace(data$age, data$age == "[80-90)", "7")
data$age <- replace(data$age, data$age == "[90-100)", "7")
```

```{r}
table(data$age)
```

```{r}
ggplot(data, aes(data$age)) + geom_bar()
data$age <- as.factor(data$age)
```

```{r}
# weight
table(data$weight)
summary(data$weight)
sum(is.na(data$weight))
```

Checking if the data is unbalanced
```{r}
ggplot(data, aes(data$weight)) + geom_bar()
data <- subset(data, select=-c(weight))
```

```{r}
#admission_type_id
table(data$admission_type_id)
```

Reducing the levels
```{r include=FALSE}
data$admission_type_id <- replace(data$admission_type_id, data$admission_type_id == 6, "5")
data$admission_type_id <- replace(data$admission_type_id, data$admission_type_id == 8, "5")
data$admission_type_id <- replace(data$admission_type_id, data$admission_type_id == 2, "1")
```

```{r}
table(data$admission_type_id)
data$admission_type_id <- as.factor(data$admission_type_id)
```

```{r}
# discharge_disposition_id 1 8 16 17 27 30 
table(data$discharge_disposition_id)
```

Reducing the levels, grouping similar discharge dispositions
```{r include=FALSE}
# Few are grouped together
data$discharge_disposition_id <- replace(data$discharge_disposition_id, data$discharge_disposition_id == 1,"group1")
data$discharge_disposition_id <- replace(data$discharge_disposition_id, data$discharge_disposition_id == 8,"group1")
data$discharge_disposition_id <- replace(data$discharge_disposition_id, data$discharge_disposition_id == 16,"group1")
data$discharge_disposition_id <- replace(data$discharge_disposition_id,
data$discharge_disposition_id == 17,"group1")
data$discharge_disposition_id <- replace(data$discharge_disposition_id, data$discharge_disposition_id == 27,"group1")
data$discharge_disposition_id <- replace(data$discharge_disposition_id,
data$discharge_disposition_id == 30, "group1")
```

```{r include=FALSE}
# 2 3 4 5 6 9 10 13 14 15 22 23 24 28 29 # Few are grouped together  
data$discharge_disposition_id <- replace(data$discharge_disposition_id, data$discharge_disposition_id == 2, "group2")
data$discharge_disposition_id <- replace(data$discharge_disposition_id, data$discharge_disposition_id == 3, "group2")
data$discharge_disposition_id <- replace(data$discharge_disposition_id, data$discharge_disposition_id == 4, "group2")
data$discharge_disposition_id <- replace(data$discharge_disposition_id, data$discharge_disposition_id == 5, "group2")
data$discharge_disposition_id <- replace(data$discharge_disposition_id, data$discharge_disposition_id == 6, "group2")
data$discharge_disposition_id <- replace(data$discharge_disposition_id, data$discharge_disposition_id == 9, "group2")
data$discharge_disposition_id <- replace(data$discharge_disposition_id, data$discharge_disposition_id == 10, "group2")
data$discharge_disposition_id <- replace(data$discharge_disposition_id, data$discharge_disposition_id == 13, "group2")
data$discharge_disposition_id <- replace(data$discharge_disposition_id, data$discharge_disposition_id == 14, "group2")
data$discharge_disposition_id <- replace(data$discharge_disposition_id, data$discharge_disposition_id == 15, "group2")
data$discharge_disposition_id <- replace(data$discharge_disposition_id, data$discharge_disposition_id == 22, "group2")
data$discharge_disposition_id <- replace(data$discharge_disposition_id, data$discharge_disposition_id == 23, "group2")
data$discharge_disposition_id <- replace(data$discharge_disposition_id, data$discharge_disposition_id == 24, "group2")
data$discharge_disposition_id <- replace(data$discharge_disposition_id, data$discharge_disposition_id == 28, "group2")
data$discharge_disposition_id <- replace(data$discharge_disposition_id, data$discharge_disposition_id == 29, "group2")
```

```{r include=FALSE}
# 7 12 18 25 26 # Few are grouped together reducing the factor levels 
data$discharge_disposition_id <- replace(data$discharge_disposition_id, data$discharge_disposition_id == 7, "group3") 
data$discharge_disposition_id <- replace(data$discharge_disposition_id, data$discharge_disposition_id == 12, "group3") 
data$discharge_disposition_id <- replace(data$discharge_disposition_id, data$discharge_disposition_id == 18, "group3") 
data$discharge_disposition_id <- replace(data$discharge_disposition_id, data$discharge_disposition_id == 25, "group3") 
data$discharge_disposition_id <- replace(data$discharge_disposition_id, data$discharge_disposition_id == 26, "group3")   
```

```{r}
# 11 19 20 21# Few are grouped together
data$discharge_disposition_id <- replace(data$discharge_disposition_id, data$discharge_disposition_id == 11, "group4")  
data$discharge_disposition_id <- replace(data$discharge_disposition_id, data$discharge_disposition_id == 19, "group4")  
data$discharge_disposition_id <- replace(data$discharge_disposition_id, data$discharge_disposition_id == 20, "group4")  
data$discharge_disposition_id <- replace(data$discharge_disposition_id, data$discharge_disposition_id == 21, "group4")  
```

```{r include=FALSE}
# group1;-60372, group2:-34436, group3:-5306, group4:-1652  # Few are grouped together
data$discharge_disposition_id <- replace(data$discharge_disposition_id, data$discharge_disposition_id == "group1", "1") 
data$discharge_disposition_id <- replace(data$discharge_disposition_id, data$discharge_disposition_id == "group2", "2") 
data$discharge_disposition_id <- replace(data$discharge_disposition_id, data$discharge_disposition_id == "group3", "3") 
data$discharge_disposition_id <- replace(data$discharge_disposition_id, data$discharge_disposition_id == "group4", "4") 
```

```{r}
table(data$discharge_disposition_id)
summary(data$discharge_disposition_id)
data$discharge_disposition_id <- as.factor(data$discharge_disposition_id)
```

```{r}
# admission_source_id
table(data$admission_source_id)
ggplot(data, aes(data$admission_source_id)) + geom_bar()
```

```{r include=FALSE}
# 1,2,3,8 # Few are grouped together
data$admission_source_id <- replace(data$admission_source_id, data$admission_source_id == 1, "group1")
data$admission_source_id <- replace(data$admission_source_id, data$admission_source_id == 2, "group1")
data$admission_source_id <- replace(data$admission_source_id, data$admission_source_id == 3, "group1")
data$admission_source_id <- replace(data$admission_source_id, data$admission_source_id == 8, "group1")

# 4, 5, 6, 7, 10,18, 19 ,22, 25, 26 # Few are grouped together
data$admission_source_id <- replace(data$admission_source_id, data$admission_source_id == 4, "group2")
data$admission_source_id <- replace(data$admission_source_id, data$admission_source_id == 5, "group2")
data$admission_source_id <- replace(data$admission_source_id, data$admission_source_id == 6, "group2")
data$admission_source_id <- replace(data$admission_source_id, data$admission_source_id == 7, "group2")
data$admission_source_id <- replace(data$admission_source_id, data$admission_source_id == 10, "group2")
data$admission_source_id <- replace(data$admission_source_id, data$admission_source_id == 18, "group2")
data$admission_source_id <- replace(data$admission_source_id, data$admission_source_id == 19, "group2")
data$admission_source_id <- replace(data$admission_source_id, data$admission_source_id == 22, "group2")
data$admission_source_id <- replace(data$admission_source_id, data$admission_source_id == 25, "group2")
data$admission_source_id <- replace(data$admission_source_id, data$admission_source_id == 26, "group2")

# 9, 15, 17, 20, 21 # Few are grouped together
data$admission_source_id <- replace(data$admission_source_id, data$admission_source_id == 9, "group3")
data$admission_source_id <- replace(data$admission_source_id, data$admission_source_id == 15, "group3")
data$admission_source_id <- replace(data$admission_source_id, data$admission_source_id == 17, "group3")
data$admission_source_id <- replace(data$admission_source_id, data$admission_source_id == 20, "group3")
data$admission_source_id <- replace(data$admission_source_id, data$admission_source_id == 21, "group3")

# 11, 12, 13, 14, 23, 24 # Few are grouped together
data$admission_source_id <- replace(data$admission_source_id, data$admission_source_id == 11, "group4")
data$admission_source_id <- replace(data$admission_source_id, data$admission_source_id == 12, "group4")
data$admission_source_id <- replace(data$admission_source_id, data$admission_source_id == 13, "group4")
data$admission_source_id <- replace(data$admission_source_id, data$admission_source_id == 14, "group4")
data$admission_source_id <- replace(data$admission_source_id, data$admission_source_id == 23, "group4")
data$admission_source_id <- replace(data$admission_source_id, data$admission_source_id == 24, "group4")
data$admission_source_id <- replace(data$admission_source_id, data$admission_source_id == "group1", "1")
data$admission_source_id <- replace(data$admission_source_id, data$admission_source_id == "group2", "2")
data$admission_source_id <- replace(data$admission_source_id, data$admission_source_id == "group3", "3")
data$admission_source_id <- replace(data$admission_source_id, data$admission_source_id == "group4", "4")
```

```{r}
table(data$admission_source_id)
data$admission_source_id <- as.factor(data$admission_source_id)
```

```{r}
# time_in_hospital
table(data$time_in_hospital)
ggplot(data, aes(data$time_in_hospital)) + geom_bar()
str(data$time_in_hospital)
```

```{r}
# Payercode
ggplot(data, aes(data$payer_code)) + geom_bar()
 #  remove unwanted variable, not helpful as not patient disease info
data <- subset(data, select=-c(payer_code))
```

```{r}
# medical_specialty
table(data$medical_specialty)
ggplot(data, aes(data$medical_specialty)) + geom_bar()
#  remove unwanted variable
data <- subset(data, select=-c(medical_specialty))
```

```{r}
ggplot(data, aes(data$num_lab_procedures)) + geom_bar()

quantile(data$num_lab_procedures, seq(0,1, 0.01))
data$num_lab_procedures[data$num_lab_procedures > quantile(data$num_lab_procedures, 0.99)] <- 
  quantile(data$num_lab_procedures, 0.99)
ggplot(data, aes(data$num_lab_procedures)) + geom_bar()
```

```{r}
ggplot(data, aes(data$num_procedures)) + geom_bar()
```

```{r}
ggplot(data, aes(data$num_medications)) + geom_bar()
quantile(data$num_medications, seq(0,1, 0.01))
data$num_medications[data$num_medications > quantile(data$num_medications, 0.99)] <- 
  quantile(data$num_medications, 0.99)
ggplot(data, aes(data$num_medications)) + geom_bar()
```

```{r}
ggplot(data, aes(data$number_outpatient)) + geom_bar()
```

```{r}
ggplot(data, aes(data$number_emergency)) + geom_bar()

ggplot(data, aes(data$number_inpatient)) + geom_bar()

ggplot(data, aes(data$number_diagnoses)) + geom_bar()
```

```{r}
ggplot(data, aes(data$max_glu_serum)) + geom_bar()
# A1Cresult compensate for this column
#  remove unwanted variable
data <- subset(data, select=-c(max_glu_serum))
#A1Cresult
ggplot(data, aes(data$A1Cresult)) + geom_bar()
ggsave("figs/A1C.png")
data$A1Cresult <- as.factor(data$A1Cresult)
```

```{r}
# remove unwanted variable as this is not taken in to account in this analysis

data <- subset(data, select=-c(metformin,repaglinide,nateglinide,chlorpropamide,glimepiride,
                        acetohexamide,glipizide,glyburide,tolbutamide,pioglitazone,
                        rosiglitazone,acarbose,miglitol,troglitazone,tolazamide,examide,
                        citoglipton,glyburide.metformin,glipizide.metformin,
                        glimepiride.pioglitazone,metformin.rosiglitazone,
                        metformin.pioglitazone))
# head(data)
```

```{r}
#change
ggplot(data, aes(data$change)) + geom_bar()
data$change <- replace(data$change, data$change == "Ch", "1")
data$change <- replace(data$change, data$change == "No", "0")
# convert to factor
summary(factor(data$change))
data$change <- as.numeric(data$change)
str(data$change)
```

```{r}
#diabetesMed
ggplot(data, aes(data$diabetesMed)) + geom_bar()
data$diabetesMed <- replace(data$diabetesMed, data$diabetesMed == "Yes", "1")
data$diabetesMed <- replace(data$diabetesMed, data$diabetesMed == "No", "0")
#convert gender to as.numeric
summary(factor(data$diabetesMed))
data$diabetesMed <- as.numeric(data$diabetesMed)
str(data$diabetesMed)
```

```{r}
#readmitted
ggplot(data, aes(data$readmitted)) + geom_bar()
data$readmitted <- replace(data$readmitted, data$readmitted == "<30", "YES")
data$readmitted <- replace(data$readmitted, data$readmitted == ">30", "YES")
ggplot(data, aes(data$readmitted)) + geom_bar()
summary(factor(data$readmitted))
```

Feature engineering

Converting comorbidity conditions to analyseable format

```{r}
## diab_code <- 250.xx
## circulatory_code <- 390-459, 785
## Not diabetic or have BP=0, diabetic alone=1, BP alone=2, both diabetic and BP=3

diagnosis <- data.frame(diag1 = as.character(data$diag_1),
                        diag2 = as.character(data$diag_2),
                        diag3 = as.character(data$diag_3),
                        stringsAsFactors = F)
str(diagnosis)
diagnosis[is.na(diagnosis) == T] <- 0

make_comorbidity <- function(x){
  
  comorbidity <- character(length = nrow(data))
  diabetes_code <- "^[2][5][0]"
  circulatory_code <- "^[3][9][0-9]|^[4][0-5][0-9]"
  
  if((str_detect(x[1], diabetes_code) == F & (str_detect(x[2], circulatory_code) == F & str_detect(x[3], circulatory_code) == F)) |
     (str_detect(x[2], diabetes_code) == F & (str_detect(x[1], circulatory_code) == F & str_detect(x[3], circulatory_code) == F)) |
     (str_detect(x[3], diabetes_code) == F & (str_detect(x[1], circulatory_code) == F & str_detect(x[2], circulatory_code) == F))){
    comorbidity <- 0
  }
  
  if((str_detect(x[1], diabetes_code) == T & (str_detect(x[2], circulatory_code) == F & str_detect(x[3], circulatory_code) == F)) |
     (str_detect(x[2], diabetes_code) == T & (str_detect(x[1], circulatory_code) == F & str_detect(x[3], circulatory_code) == F)) |
     (str_detect(x[3], diabetes_code) == T & (str_detect(x[1], circulatory_code) == F & str_detect(x[2], circulatory_code) == F))){
    comorbidity <- 1
  }
  
  if((str_detect(x[1], circulatory_code) == T & (str_detect(x[2], diabetes_code) == F & str_detect(x[3], diabetes_code) == F)) |
     (str_detect(x[2], circulatory_code) == T & (str_detect(x[1], diabetes_code) == F & str_detect(x[3], diabetes_code) == F)) |
     (str_detect(x[3], circulatory_code) == T & (str_detect(x[1], diabetes_code) == F & str_detect(x[2], diabetes_code) == F))){
    comorbidity <- 2
  }
  
  if((str_detect(x[1], diabetes_code) == T & (str_detect(x[2], circulatory_code) == T | str_detect(x[3], circulatory_code) == T)) |
     (str_detect(x[2], diabetes_code) == T & (str_detect(x[1], circulatory_code) == T | str_detect(x[3], circulatory_code) == T)) |
     (str_detect(x[3], diabetes_code) == T & (str_detect(x[1], circulatory_code) == T | str_detect(x[2], circulatory_code) == T))){
    comorbidity <- 3
  }
  comorbidity <- ordered(comorbidity, levels = c(0,1,2,3))
  return(comorbidity)
}
data$comorbidity <- apply(diagnosis, 1, make_comorbidity)

# remove diagnosis 
data$diag_1 <- NULL
data$diag_2 <- NULL
data$diag_3 <- NULL
```

```{r}
# comorbidity and readmission 
prop.table(table(data$readmitted, data$comorbidity), margin = 2)
```

```{r}
# analyse comorbidity
ggplot(data, aes(x=comorbidity, y=number_inpatient)) + geom_bar(stat = "identity")

ggplot(data, aes(x=comorbidity, y=number_outpatient)) + geom_bar(stat = "identity")

ggplot(data, aes(x=comorbidity, y=number_emergency)) + geom_bar(stat = "identity")

tapply(data$number_inpatient, data$comorbidity, mean)
tapply(data$number_emergency, data$comorbidity, mean)
tapply(data$number_outpatient, data$comorbidity, mean)
tapply(data$time_in_hospital, data$comorbidity, mean)
tapply(data$number_diagnoses, data$comorbidity, mean)
tapply(data$num_lab_procedures, data$comorbidity, mean)
tapply(data$num_procedures, data$comorbidity, mean)
tapply(data$num_medications, data$comorbidity, mean)
```


```{r}
# Analyse A1Cresult
ggplot(data, aes(x=A1Cresult, y=number_inpatient)) + geom_bar(stat = "identity")

ggplot(data, aes(x=A1Cresult, y=number_outpatient)) + geom_bar(stat = "identity")

ggplot(data, aes(x=A1Cresult, y=number_emergency)) + geom_bar(stat = "identity")

tapply(data$number_inpatient, data$A1Cresult, mean)
tapply(data$number_outpatient, data$A1Cresult, mean)
tapply(data$number_emergency, data$A1Cresult, mean)
```

```{r}
#Missing values
sapply(data, function(x) sum(is.na(x)))
#Percentage of missing values
colMeans(is.na(data))
```

```{r}
# Removing NAs
# before removing NAs patient ID is joined to the data
data <- cbind(Patient_Id, data)
data<-na.omit(data)
ID <- data$Patient_Id
data <- data[, 2:21]
```

Dummay variables are created
```{r}

# race
#creating dummy variable for race variable    
dummy_race <- model.matrix(~race - 1,data=data)
# Removing the 1st dummy variable
dummy_race <- dummy_race[,-1]
# Adding dummy variable of race to data and removing race column
data <- cbind(data[,-1], dummy_race)

#age
#creating dummy variable for age variable    
dummy_age <- model.matrix(~age - 1,data=data)
# Removing the 1st dummy variable
dummy_age <- dummy_age[,-1]
# Adding dummy variable of age to data and removing age column
data <- cbind(data[,-2], dummy_age)

#admission_type_id
#creating dummy variable for admission_type_id variable    
dummy_admission_type_id <- model.matrix(~admission_type_id - 1,data=data)
# Removing the 1st dummy variable
dummy_admission_type_id <- dummy_admission_type_id[,-1]
# Adding dummy variable of admission_type_id to data and removing admission_type_id column
data <- cbind(data[,-2], dummy_admission_type_id)

#discharge_disposition_id
#creating dummy variable for discharge_disposition_id variable    
dummy_discharge_disposition_id <- model.matrix(~discharge_disposition_id - 1,data=data)
# Removing the 1st dummy variable
dummy_discharge_disposition_id <- dummy_discharge_disposition_id[,-1]
# Adding dummy variable of discharge_disposition_id to data and removing discharge_disposition_id column
data <- cbind(data[,-2], dummy_discharge_disposition_id)

#admission_source_id
#creating dummy variable for admission_source_id variable    
dummy_admission_source_id <- model.matrix(~admission_source_id - 1,data=data)
# Removing the 1st dummy variable
dummy_admission_source_id <- dummy_admission_source_id[,-1]
# Adding dummy variable of admission_source_id to data and removing admission_source_id column
data <- cbind(data[,-2], dummy_admission_source_id)

#A1Cresult
#creating dummy variable for A1Cresult variable    
dummy_A1Cresult <- model.matrix(~A1Cresult - 1,data=data)
# Removing the 1st dummy variable
dummy_A1Cresult <- dummy_A1Cresult[,-1]
# Adding dummy variable of A1Cresult to data and removing A1Cresult column
data <- cbind(data[,-10], dummy_A1Cresult)

#insulin
#creating dummy variable for insulin variable    
dummy_insulin <- model.matrix(~insulin - 1,data=data)
# Removing the 1st dummy variable
dummy_insulin <- dummy_insulin[,-1]
# Adding dummy variable of insulin to data and removing insulin column
data <- cbind(data[,-10], dummy_insulin)

#comorbidity
#creating dummy variable for comorbidity variable    
dummy_comorbidity <- model.matrix(~comorbidity - 1,data=data)
# Removing the 1st dummy variable
dummy_comorbidity <- dummy_comorbidity[,-1]
# Adding dummy variable of race to data and removing race column
data <- cbind(data[,-13], dummy_comorbidity)
```

```{r include=FALSE}
# Feature standardisation
# Normalising continuous features, scaling
data$time_in_hospital <- scale(data$time_in_hospital)
data$num_lab_procedures <- scale(data$num_lab_procedures)
data$num_procedures <- scale(data$num_procedures)
data$num_medications <- scale(data$num_medications)
data$number_outpatient <- scale(data$number_outpatient)
data$number_emergency <- scale(data$number_emergency)
data$number_inpatient <- scale(data$number_inpatient)
data$number_diagnoses <- scale(data$number_diagnoses)
```

```{r}
# Readmitted is separated and added to the first column in the dataset and removed the older column
readmitted <- data$readmitted
data <- cbind(readmitted, data)
data <- data[,-13]
# All are numerical variable.
```

```{r}
# join the ID of patients
data <- cbind(ID, data)
str(data)
```
